We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-06 paper/paper.tex
abstract

\begin{abstract}
We use numerical simulations to build an explicit probability distribution for the
asphericity in the satellite distribution around galaxies similar to the
Local Group (LG) in the Lambda Cold Dark Matter (LCDM) paradigm.
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ greatly among the simulations, suggesting a
strong influence of resolution, simulated physics and criteria to select
LG pairs.
However, there are three common trends in these simulations. First, at most of
$1\%$ the pairs are expected to have satellite distributions with the
same asphericity as the Local Group; second, between
$38\%$ to $59\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $2\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in M31 galaxy turn out to have a rather typical distribution 
in the LCDM context, while the MW is atypical in this respect. 
These results place the LG close to the level of a 3-$\sigma$ outlier
in the LCDM paradigm.
The approach presented here will facilitate the comparison between
asphericity results from different numerical setups and choices to
select halo pairs in simulations, hopefully helping to clarify the
reasons behin the LG atipicallity in LCDM. 
\end{abstract}

>>>>

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ greatly among the simulations, suggesting a
strong influence of resolution, simulated physics and criteria to select
LG pairs.
However, there are three common trends in these simulations. 
First, at most of $1\%$ the pairs are expected to have satellite
distributions with the same asphericity aas the Local Group; second,
between $38\%$ to $59\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $2\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in M31 galaxy turn out to have a rather
typical distribution in the LCDM context, while the MW is atypical in
this respect.  
These results place the LG close to the level of a 3-$\sigma$ outlier
in the LCDM paradigm. 
The approach presented here facilitates the comparison between
asphericity results from different numerical setups and choices to
select halo pairs in simulations, hopefully helping to clarify the
reasons behin the LG atipicallity in LCDM. 
\end{abstract}
